---
title: "Outliersnps"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Population analyses on Outlier dataset  


```{r}
#Loading all the necessary packages
library(adegenet)
library(vcfR)
library("radiator") # Conversion from vcf to a lot of other formats
library("dplyr")
library("hierfstat")
library("ggplot2") #For plotting
library("reshape2") #For plotting
library("plyr")
library("cowplot") #For plotting manuscript figs
library(PCAviz)  #Visualizing output of PCA
```


### Making files


#### Make genind object
```{r}
my_vcf <- read.vcfR("outliersnps.recode.vcf")

strata<- read.table("strata", header=TRUE)
```

```{r}
rad.filt <- vcfR2genind(my_vcf, strata = strata, pop = c(rep("FBN", 30),rep("FBS", 25),rep("OBN", 19), rep("OBS",28), rep("PCN",27), rep("PCS",25), rep("SPN",28), rep("SPS",29), rep("WC1",25), rep("WC2",27), rep("WC3",28), rep("WC4",29)))
```

```{r}
rad.filt
```

```{r}
#Providing population names for plotting
pop_order <- c("FBN","FBS","OBN","OBS","PCN","PCS","SPN","SPS","WC1","WC2","WC3","WC4")
```

#### Read in the other info from .strata file and extract information such as locality, latitude, and longitude.
```{r}
info <- as.data.frame(read.table("strata",header = T,sep = "\t",stringsAsFactors = F))

mystrats <- as.data.frame(matrix(,nrow = length(indNames(rad.filt)),ncol=5))
colnames(mystrats) <- c("Population","Locality","Latitude","Longitude","Distance")

just.strats <- select(info,c("Population","Locality"))
stratted.filt <- strata(rad.filt, formula= Locality/Population, combine = TRUE,just.strats)
stratted.filt@other <- select(info, Latitude,Longitude,Distance)
```

```{r}
stratted.filt
```

### Make hierfstat object

```{r}
hf.filt <- genind2hierfstat(rad.filt, pop = c(rep("FBN", 30),rep("FBS", 25),rep("OBN", 19), rep("OBS",28), rep("PCN",27), rep("PCS",25), rep("SPN",28), rep("SPS",29), rep("WC1",25), rep("WC2",27), rep("WC3",28), rep("WC4",29))) 
```

```{r}
hf.u <- genind2hierfstat(rad.u, pop = c(rep("FBN", 30),rep("FBS", 25),rep("OBN", 19), rep("OBS",28), rep("PCN",27), rep("PCS",25), rep("SPN",28), rep("SPS",29), rep("WC1",25), rep("WC2",27), rep("WC3",28), rep("WC4",29)))
hf.u <- hf.u$hierfstat.no.imputation
```


## Pairwise Fst

```{r}
fst.mat <- pairwise.WCfst(hf.filt)
```


```{r}
gindF.fst.mat.triN <- as.matrix(fst.mat)
colnames(gindF.fst.mat.triN) <- pop_order
rownames(gindF.fst.mat.triN) <- pop_order
```

```{r}
meltedN <- melt(gindF.fst.mat.triN, na.rm =TRUE)
round(gindF.fst.mat.triN,4)
```

```{r}
summary(meltedN$value)
```


```{r}
#Plotting Pairwise fst
neutral <- ggplot(data = meltedN, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "darkorange2", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984) Neutral", atop(italic("N = 320, L = 118"), ""))))+
  labs( x = "Sampling Site", y = "Sampling Site") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 13)) + 
  theme(axis.title = element_text(size = 16),legend.text = element_text(size =15), legend.title = element_text(size =16)) +
  theme(plot.title = element_text(size = 17)) +
coord_fixed()
neutral
```

### Re-running Pairwise Fst using regions instead of populations
```{r}
# make a copy of the hierfstat object
hf.reg <- hf.filt
# relabel individuals by their phylogeographic region
regions <- c("FB","FB","OB","OB","PC","PC","SP","SP",rep("WC",4))
hf.reg$pop <- mapvalues(hf.reg$pop, from = c("FBN","FBS","OBN","OBS","PCN","PCS","SPN","SPS","WC1","WC2","WC3","WC4"), to = regions)
```

```{r}
fst.mat.reg.N <- pairwise.WCfst(hf.reg)
```

```{r}
fst.mat.reg.N
```

```{r}
NS <- c("FB","OB","PC","SP","WC")
gindF.fst.mat.triR <- as.matrix(fst.mat.reg.N)
gindF.fst.mat.triR <- gindF.fst.mat.triR[NS,NS]
```

```{r}
gindF.fst.mat.triR
```

```{r}
meltedN_reg <- melt(gindF.fst.mat.triR, na.rm =TRUE)
```

```{r}
#Plotting Pairwise fst
neut_reg <- ggplot(data = meltedN_reg, aes(Var2, Var1, fill = value))+ geom_tile(color = "white")+ 
  scale_fill_gradient(low = "white", high = "red", name="FST")  +
  ggtitle(expression(atop("Pairwise FST, WC (1984), Regions", atop(italic("N = 320, L = 6663"), ""))))+
  labs( x = "Region", y = "Region") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),axis.text.y = element_text(size = 13)) + 
  coord_fixed()
neut_reg
```


## Genetic diversity (observed and expected heterozygosity)

```{r}
comb <- summary(stratted.filt)
names(comb)
```

```{r}
plot(comb$Hobs, xlab="Loci number", ylab="Observed Heterozygosity", 
     main="Observed heterozygosity per locus")
```


```{r}
plot(comb$Hobs,comb$Hexp, xlab="Hobs", ylab="Hexp", 
     main="Expected heterozygosity as a function of observed heterozygosity per locus")
```


```{r}
bartlett.test(list(comb$Hexp, comb$Hobs)) # a test : H0: Hexp = Hobs
```


#### Not a significant difference between Observed and expected heterzygosity 


```{r}
basicstat <- basic.stats(hf.filt, diploid = TRUE, digits = 3)
```

```{r}
as.data.frame(basicstat$overall)
```

```{r}
# get bootstrap confidence values for Fis
boot <- boot.ppfis(hf.filt,nboot = 1000)
boot5 <- boot.ppfis(hf.filt,nboot = 1000,quant = 0.5)
```

```{r}
# add latitude for each population
latitude = c(27.674833, 27.667618, 27.708808, 27.696972, 27.632041, 27.614554, 27.564633, 27.473143, 27.589457, 27.59879, 27.605984, 27.614192)
```

```{r}
# add longitude for each population
longitude = c(-97.268409, -97.272238, -97.336006, -97.331832, -97.213456, -97.213456, -97.257698, -97.311179, -97.261131, -97.256947, -97.252883, -97.248581)
```

```{r}
# add distance for each population
distance = c(0.617, 0.27, 0.22, 1.555, 6.16, 7.77, 4.077, 15.358, 1.547, 0.516, 0.56, 1.516)
```

```{r}
# combine all pop statistics
colnames(basicstat$Ho) <- pop_order
Ho <- colMeans(basicstat$Ho,na.rm = T)
He <- colMeans(basicstat$Hs,na.rm = T)
Fis<- boot5$fis.ci$ll
y <- cbind(Ho,He,Fis,boot$fis.ci,latitude,longitude,distance)
y
```

```{r}
summary(He)
```

```{r}
summary(Fis)
```

```{r}
# Plot He vs Latitude
R2 = round(summary(lm(y$He ~ y$latitude))$r.squared, 4)             

ggplot(y, aes(x = latitude, y = He)) + geom_point() + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Latitude, Neutral") +
annotate(geom = "text", label=paste("italic(R^2)==",R2), x=27.7, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
# Plot He vs Longitude
R3 = round(summary(lm(y$He ~ y$longitude))$r.squared, 4)             

ggplot(y, aes(x = longitude, y = He)) + geom_point() + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Longitude, Neutral") +
annotate(geom = "text", label=paste("italic(R^2)==",R3), x=-97.22, y=0.2635, parse=T) +
scale_x_reverse()
```

```{r}
# Plot He vs Distance 
R4 = round(summary(lm(y$He ~ y$distance))$r.squared, 4)             

ggplot(y, aes(x = distance, y = He)) + geom_point() + 
geom_smooth(method=lm) + 
ggtitle("Expected heterozygosity vs Distance, Neutral") +
annotate(geom = "text", label=paste("italic(R^2)==",R4), x=0, y=0.2635, parse=T) +
scale_x_reverse()
```


## PCA 

```{r}
X <- tab(stratted.filt, freq = TRUE, NA.method = "mean")
pca1 <- dudi.pca(X, scale = FALSE, scannf = FALSE, nf = 3)
barplot(pca1$eig[1:50], main = "PCA eigenvalues", col = heat.colors(50))
s.class(pca1$li, pop(stratted.filt))
title("PCA of simulated dataset\naxes 1-2")
add.scatter.eig(pca1$eig[1:20], 3,1,2)

col <- funky(15) 
s.class(pca1$li, pop(stratted.filt),xax=1,yax=2, col=col, axesell=FALSE, cstar=0, cpoint=3, grid=FALSE)
```

## DAPC

```{r}
grp <- (stratted.filt@pop)
```

```{r}
table(pop(stratted.filt), grp)

table.value(table(pop(stratted.filt), grp), col.lab=paste("inf", 1:2), row.lab=paste("ori", 1:4))
```

```{r}
dapc1 <- dapc(stratted.filt, grp)
```

```{r}
scatter(dapc1,col=col,bg="white", solid=1)
```

```{r}
#Structure Like
compoplot(dapc1, posi="bottomright",txt.leg=paste("Cluster", 1:12), lab="", ncol=1, xlab="individuals")

temp <- which(apply(dapc1$posterior,1, function(e) all(e<0.9)))

compoplot(dapc1, subset=temp, posi="bottomright", txt.leg=paste("Cluster", 1:12), ncol=2)
```


## PCAviz

```{r}
NA.afDraw<- function(ind){
  ind.mat <- ind@tab
  new.mat <- ind.mat
  af = colSums(ind.mat[,seq(1,ncol(ind.mat)-1,2)],na.rm = TRUE)/
      (2*apply(ind.mat[,seq(1,ncol(ind.mat)-1,2)],2,function(x) sum(!is.na(x))))
  af.Draw <- function(geno, af){
     new <- function(geno,af){
        if(is.na(geno)){
        newA = rbinom(1,2,af)
        }
        else {newA <- geno}
        return(newA)
   }
  new.row <- mapply(geno,af,FUN = new)
  return(new.row)}
  
  new.mat[,seq(1,ncol(ind.mat)-1,2)] <- t(apply(ind.mat[,seq(1,ncol(ind.mat)-1,2)],1,af.Draw,af))
  new.mat[,seq(2,ncol(ind.mat),2)] <- 2-new.mat[,seq(1,ncol(ind.mat)-1,2)]
  new.ind <- ind
  new.ind@tab <- new.mat
  return(new.ind)
}
```

```{r}
u.na <- NA.afDraw(stratted.filt)
```

```{r}
pca <- dudi.pca(u.na,cent=TRUE,scale=TRUE)
```

```{r}
pca <- dudi.pca(u.na,cent=TRUE,scale=TRUE,scannf = F, nf = 30)
```

```{r}
col18 <- funky(length(unique(u.na@strata$Population)))
#Colors that match the neutral Structure results
col6 <-  c("firebrick","dodgerblue","midnightblue","salmon","forestgreen","chocolate1")

par(mfrow=c(2,2))
s.class(pca$li, strata(u.na)$Locality,xax=1,yax=2, 
        sub = "neutralloci,PC 1-2, 320 individuals, 6,663 SNPs, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
s.class(pca$li, strata(u.na)$Locality,xax=1,yax=3, 
        sub = "neutralloci,PC 1-2, 320 individuals, 6,663 SNPs, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
s.class(pca$li, strata(u.na)$Locality,xax=1,yax=4, 
        sub = "neutralloci,PC 1-2, 320 individuals, 6,663 SNPs, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
s.class(pca$li, strata(u.na)$Locality,xax=1,yax=5, 
        sub = "neutralloci,PC 1-2, 320 individuals, 6,663 SNPs, maf 5.0%",
        possub = "topleft",col=transp(col6,.6),axesell=FALSE,
        cstar=0, cpoint=3, grid=FALSE, cellipse = 0)
```

```{r}
s.label(pca$li, xax=1,yax=5, 
        sub = "neutralloci,PC 1-2, 320 individuals, 6,663 SNPs, maf 5.0%",
        possub = "topleft")
```


```{r}
eig.perc <- 100*pca$eig/sum(pca$eig)
head(eig.perc)
```

```{r}
li <-pca$li
c1 <- pca$c1
#Create dataframe of info like latitude and population for each individual
info_mat <- as.data.frame(cbind(u.na@strata, u.na@other$Latitude, u.na@other$Longitude, u.na@other$Distance))
colnames(info_mat) <- c("Population","Locality","Latitude", "Longitude", "Distance")
colnames(c1) <- colnames(li)
#create pcaviz object
pviz <- pcaviz(x=li,rotation=c1,dat=info_mat)
```


```{r}
#Comparing to Distance
p = list(size=2) 
plot(pviz,color = "Distance", draw.points = T, group.summary.labels = TRUE, draw.pc.axes = T, geom.point.params = p)
```

```{r}
p = list(size=6)
plot(pviz, coords = c("Axis1","Distance"),
     show.legend = T,color = "Population",colors = col18,
     draw.points =T,group.summary.labels = F,geom.point.params= list(size = 6, alpha = 0.9),geom.point.summary.params = list(size =6,alpha = 0.4))
```

```{r}
#Comparing to Latitude
p = list(size=2) 
plot(pviz,color = "Latitude", draw.points = T, group.summary.labels = TRUE, draw.pc.axes = T, geom.point.params = p)
```

```{r}
p = list(size=2) 
plot(pviz,color = "Longitude", draw.points = T, group.summary.labels = TRUE, draw.pc.axes = T, geom.point.params = p)
```

```{r}
p = list(size=6)
plot(pviz, coords = c("Axis1","Latitude"),
     show.legend = T,color = "Population",colors = col18,
     draw.points =T,group.summary.labels = F,geom.point.params= list(size = 6, alpha = 0.9),geom.point.summary.params = list(size =6,alpha = 0.4))
```

```{r}
plot(pviz,coords = c("Axis1","Axis2","Axis3","Axis4","Axis5","Distance"),group = NULL,color= "Population",colors = col18,draw.points = T,scale.pc.axes = 0.6, show.legend=F)
```

```{r}
p = list(size=6)
A2 <- plot(pviz,coords = c("Axis2","Distance"),group="Population",
     show.legend = F,color = "Locality",colors = col6,draw.points =T,
     group.summary.labels = F,draw.linear.fit = F, coord_fixed=T)
A3 <- plot(pviz,coords = c("Axis3","Distance"),group="Population",
     show.legend = F,color = "Locality",colors = col6,
     draw.points =T,group.summary.labels = F,draw.linear.fit = F)
A4 <- plot(pviz,coords = c("Axis4","Distance"),group="Population",
     show.legend = F,color = "Locality",colors = col6,
     draw.points =T,group.summary.labels = F,draw.linear.fit = F)
A5 <- plot(pviz,coords = c("Axis5","Distance"),group="Population",
     show.legend = T,color = "Locality",colors = col6,
     draw.points =T,group.summary.labels = F,draw.linear.fit = F)
plot_grid(A2,A3,A4,A5, labels = c('A','B','C','D'))
```